---
title: "Untitled"
output: pdf_document
---

```

# Raw log2-transformation
To compare with other normalization methods.

```{r raw log2-transformation}
assay(sce_endo, "logcounts_raw") <- log2(counts(sce_endo) + 1)
# plotReducedDim and plotPCA will do the same, with plotPCA you leave out the
# use_dimred="PCA" argument.
tmp <- runPCA(sce_endo, 
              ncomponents = 50, 
              exprs_values = "logcounts_raw")
# plot PCA after log2 transformation
plotPCA(tmp,
        colour_by = params$lab_col,
        size_by = "total_features_by_counts")
# One can also run tSNE in similar ways with Scater.
rm(tmp)
# The logcounts_raw is not enough to account for the technical factors between
# the cells.
```

## Normalization in Seurat
Make the seurat object, 'seuset'. In this step you could filter the cells
again, these however already have been filtered before in 'table clean-up',
where the genes were taken that have >2 cells that have an expression >1.

```{r}
# looking into the dataset
VlnPlot(
    object = seuset,
    features = c("nFeature_sf"),
    group.by = params$lab_col
)
VlnPlot(
    object = seuset,
    features = c("nCount_sf"),
    group.by = params$lab_col
)
#VlnPlot(
#  object = seuset,
#    features = explore_violin,
#    group.by = params$lab_col
#)
FeatureScatter(
    object = seuset,
    feature1 = "nCount_sf",
    feature2 = "nFeature_sf"
)
```

```{r}
# Seurat normalization: "a global-scaling normalization method LogNormalize that
# normalizes the gene expression measurements for each cell by the total
# expression, multiplies this by a scale factor (10,000 by default), and
# log-transforms the result.""
seu <- seuset
seuset <- NormalizeData(
    object = seuset,
    normalization.method = "LogNormalize",
    scale.factor = 10000
)
# looking into the dataset
VlnPlot(
    object = seuset,
    features = c("nFeature_sf"),
    group.by = params$lab_col
)
VlnPlot(
    object = seuset,
    features = c("nCount_sf"),
    group.by = params$lab_col
)
#VlnPlot(
#  object = seuset,
#    features = explore_violin,
#    group.by = params$lab_col
#)
FeatureScatter(
    object = seuset,
    feature1 = "nCount_sf",
    feature2 = "nFeature_sf"
)
saveRDS(seuset, paste(params$resultsdir,"seuset_qc+norm.rds",sep="/"))
```

#### Check confounders before & after normalization

```{r seurat objects to sce}
# Only take the entries that are matchable with the counttable entries:
filtered_cells <- intersect(rownames(phenodata), 
                            colnames(seuset@assays$sf@data))
pheno_matchedseuset <- phenodata[filtered_cells, ]
pheno_orderedseuset <-
pheno_matchedseuset[match(colnames(seuset@assays$sf@data),
rownames(pheno_matchedseuset)), ]
count_matrixseuset <- as.matrix(seuset@assays$sf@data)
sce_seunorm <-
SingleCellExperiment(
assays = list(counts = count_matrixseuset),
colData = pheno_orderedseuset,
rowData = rownames(count_matrixseuset)
)
# A little trick to let scater know that there are actually logcounts in the dataset.
assay(sce_seunorm, "logcounts") <- counts(sce_seunorm)
# Calculate the quality metrics:
sce_seunorm <- calculateQCMetrics(
  sce_seunorm)
```

# Identifying the variation caused by each confounding factor
#### Before & after normalization

```{r check confounders in raw dataset}
explanatory_variables <- as.factor(c(confounders_to_test, 
                                     "total_features_by_counts", "total_counts"))
#explanatory_variables_seu <- c("total_features_by_counts", "total_counts",
#confounders_to_test)
# This function and visualization performs a PCA analysis in the data object and
# checks to what extend the variables that are put in, are explaining the
# variance. The percentage of variance explained by each variable of interest:
# Setting the colours:
colourvector <- c()
colourset <- brewer.pal(length(explanatory_variables),"Dark2")
i <- 1
for (variable_item in explanatory_variables){
  colourvector[variable_item] <- colourset[i]
  i <- i + 1
}
# Building combined plot, before and after normalization
p1 <- plotExplanatoryVariables(sce_endo,
                               exprs_values = "counts",
                               variables = explanatory_variables) + 
                               expand_limits(y = 1) + 
                               scale_color_manual(values = colourvector) + 
                               ggtitle("Explanatory Variables Before Normalization")
p2 <- plotExplanatoryVariables(sce_seunorm,
                               variables = explanatory_variables) + 
                               expand_limits(y = 1) + 
                               scale_color_manual(values = colourvector) + 
                               ggtitle("Explanatory Variables After Normalization")
multiplot(p1, p2)
```


```{r}
# running PCA on the normalized counts
sce_seunorm <- runPCA(
  sce_seunorm, ncomponents = 20,
  exprs_values = "counts"
)
```


```{r,  linewidth=60}
# plotting again the PCA's on raw-transformed and normalized values
# raw log-transformation.
tmp <- runPCA(sce_endo, ncomponents = 50, exprs_values = "logcounts_raw")
# PCA plot after log2 transformation
plotPCA(tmp,
        colour_by = params$lab_col,
        size_by = "total_features_by_counts")
# PCA plot after seurat normalization
plotPCA(sce_seunorm,
        colour_by = params$lab_col,
        size_by = "total_features_by_counts")
```


## Build unspliced assay

Select the same cells and genes as in the spliced dataset

```{r build SCE 2}
# df -> matrix -> SCE + phenodata
cells_use <- colnames(sce_endo)
genes_use <- rownames(sce_endo)
sce_us <-
  SingleCellExperiment(
  assays = list(counts = unspliced.data.df),
  colData = pheno_matched,
  rowData = rownames(unspliced.data.df)
  )
control_features_us <- vector("list", 0)
control_features_us_match <- vector("list", 0)
# Dataset after filtering:
sce_usmatch <- sce_us[genes_use,cells_use]
# Adding spike-in information:
if (params$add.spikes.mt) {
  isSpike(sce_us, "MT") <- grepl("^MT-", rownames(sce_us))
  control_features_us[["MT"]] <- isSpike(sce_us, "MT")
  #matched sce object
  isSpike(sce_usmatch, "MT") <- grepl("^MT-", rownames(sce_usmatch))
  control_features_us_match[["MT"]] <- isSpike(sce_usmatch, "MT")
}
if (params$add.spikes.ercc) {
  isSpike(sce_us, "ERCC") <- grepl("^ERCC-", rownames(sce_us))
  control_features_us[["ERCC"]] <- isSpike(sce_us, "ERCC")
  #matched sce object
  isSpike(sce_usmatch, "ERCC") <- grepl("^ERCC-", rownames(sce_usmatch))
  control_features_us_match[["ERCC"]] <- isSpike(sce_usmatch, "ERCC")
}
# Calculate the quality metrics:
# Calculate the quality metrics:
sce_us <- calculateQCMetrics(
  sce_us, feature_controls = control_features_us
    )
sce_usmatch <- calculateQCMetrics(
  sce_usmatch, feature_controls = control_features_us_match
  )
# Arbitrary thresholds:
# Looking at the total number of RNA molecules per sample
# UMI counts were used for this experiment
hist(sce_us$total_counts, breaks = 100)
abline(v = params$total_counts_tresh, col = "red")
# Looking at the amount of unique genes per sample
# This is the amount with ERCC included.
hist(sce_us$total_features_by_counts, breaks = 100)
abline(v= params$total_feat_tresh, col = "red")
hist(sce_usmatch$total_counts, breaks = 100)
abline(v = params$total_counts_tresh, col = "red")
hist(sce_usmatch$total_features_by_counts, breaks = 100)
abline(v= params$total_feat_tresh, col = "red")
pdf(
  paste(
  params$resultsdir,
  "Histograms_before+aftercellsFiltering_UnsplicedReads.pdf",
  sep = "/"
  )
  )
par(mfrow=c(2,2))
hist(sce_us$total_counts, breaks = 100)
abline(v = params$total_counts_tresh, col = "red")
hist(sce_us$total_features_by_counts, breaks = 100)
abline(v= params$total_feat_tresh, col = "red")
hist(sce_usmatch$total_counts, breaks = 100)
abline(v = params$total_counts_tresh, col = "red")
hist(sce_usmatch$total_features_by_counts, breaks = 100)
abline(v= params$total_feat_tresh, col = "red")
dev.off()
```
## Build Seurat object with unspliced and spliced assay

```{r}
unspliced_match <- unspliced.data.df[genes_use,cells_use]
unspliced_match <- as.matrix(unspliced_match)
seu[["uf"]] <- CreateAssayObject(counts = unspliced_match)
seu <- NormalizeData(
    object = seu, assay = "sf",
    normalization.method = "LogNormalize",
    scale.factor = 10000
)
seu <- NormalizeData(
    object = seu, assay = "uf",
    normalization.method = "LogNormalize",
    scale.factor = 10000
)
```


## Highly variable genes & Scaling of the gene expression values

```{r}
# FindVariableFeatures plots the dispersion (= a normalized measure of
# cell-to-cell variation), as a function of average expression for each gene. In
# their tutorial the Satija lab uses the cut-off of 2000 genes.
seu <- FindVariableFeatures(
    object = seu, assay = "sf",
    selection.method = "vst",
    nfeatures = params$nhvg)
seu <- FindVariableFeatures(
    object = seu, assay = "uf",
    selection.method = "vst",
    nfeatures = params$nhvg)
# top 10 most variable genes
top20 <- head(VariableFeatures(seu, assay = "sf"), 20)
top20_uf <- head(VariableFeatures(seu, assay = "uf"), 20)
# plot variable features with labels:
plot1 <- VariableFeaturePlot(seu)
plot2 <- LabelPoints(plot = plot1, points = top20, repel = TRUE)
plot2
plot3 <- VariableFeaturePlot(seu, assay = "uf")
plot4 <- LabelPoints(plot = plot3, points = top20_uf, repel = TRUE)
plot4
# Preferable removing the genes that are highly expressed but with a low variance.
length(x = seu@assays$sf@var.features)
seu[["sf"]]@var.features[1:10]
```

```{r scaling and regressing,  linewidth=60}
# Scaling the data to make it usable for dimensional reduction
# using all the genes, could also select only the highly variable genes.
# Optional regression is performed here.
all.genes <- rownames(seuset)
seu <- ScaleData(
    object = seu,  vars.to.regress = vars_to_regress_sf,
    assay = "sf",
    features = all.genes
)
seu <- ScaleData(
    object = seu,  vars.to.regress =  vars_to_regress_uf,
    assay = "uf",
    features = all.genes
)
```

## Running PCA analysis on the scaled data
```{r running PCA}
seuset <- seu
rm(seu)
DefaultAssay(seuset) <- "sf"
seuset <- RunPCA(
    object = seuset,
    features = VariableFeatures(object = seuset),
    npcs = params$pcs_max_hvg,
    ndims.print = 1:5,
    nfeatures.print = 5
)
length(seuset[["sf"]]@var.features)
length(seuset[["uf"]]@var.features)
```
